### 2022-01-17 latexDraft/appendix/appendix.tex
ok, let's fix the bug...

\documentclass[]{article}
\usepackage{amsmath,amssymb}
\usepackage{graphicx} % to insert pictures
\usepackage{bm}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

\begin{document}

\section{$x_t^{(j)}$ and $\mu_t^{(j)}$}

Let the $t^{\text{th}}$ column observation and firing rate of the $j^{\text{th}}$ cluster be $\widetilde{\bm{y}}_t = vec\left(\{y_{it}^{(j)}|z_i = j\}\right)$ and $\widetilde{\bm{\lambda}}_t^{(j)} = vec\left(\{\lambda_{it}|z_i = j\}\right)$. The number of neurons in cluster $j$ is $n_j=\left|\{i:z_i = j\}\right|$. The corresponding loading for these $n_j$ neurons is $\bm{C}^{(j)}\in\mathbb{R}^{n_j\times p}$, such that $\log \widetilde{\bm{\lambda}}_t^{(j)} = \mu_t^{(j)}\bm{1}_{n_j} + \bm{C}^{(j)}\bm{x}_t^{(j)} = \left(\bm{1}_{n_j}, \bm{C}^{(j)} \right)\cdot \left(\mu_t^{(j)}, \bm{x}_t^{'(j)}\right)'$. Denote $\bm{\widetilde{C}}^{(j)} = \left(\bm{1}_{n_j}, \bm{C}^{(j)} \right)$, $\bm{\widetilde{x}}_t^{(j)} = \left(\mu_t^{(j)}, \bm{x}_t^{'(j)}\right)'$, $\bm{\widetilde{x}}^{(j)} = \left(\bm{\widetilde{x}}_1^{'(j)},\ldots,\bm{\widetilde{x}}_T^{'(j)}\right)'$, $\widetilde{\bm{A}}^{(j)} = diag(f^{(j)}, \bm{A}^{(j)})$, $\widetilde{\bm{b}}^{(j)} = \left(g^{(j)}, \bm{b}^{'(j)}\right)'$ and $\widetilde{\bm{Q}}^{(j)} = diag(\Sigma^{(j)}, \bm{Q}^{(j)})$. The full conditional distribution $P(\bm{\widetilde{x}}^{(j)}|\ldots) = P(\bm{\widetilde{x}}^{(j)}|\{\widetilde{\bm{y}}_t\}_{t=1}^T, \bm{\widetilde{C}}^{(j)}, \widetilde{\bm{A}}^{(j)}, \widetilde{\bm{b}}^{(j)}, \widetilde{\bm{Q}}^{(j)})$ is approximated by a global Laplace approximation, i.e. $P(\bm{\widetilde{x}}^{(j)}|\ldots) \approx  N_{(p+1)T}(\bm{\widetilde{x}}^{(j)} | \bm{\mu}_{\widetilde{\bm{x}}^{(j)}}, \bm{\Sigma}_{\widetilde{\bm{x}}^{(j)}})$, with $\bm{\mu}_{\widetilde{\bm{x}}^{(j)}} = \argmax_{\widetilde{\bm{x}}^{(j)}}P(\bm{\widetilde{x}}^{(j)}|\ldots)$ and $\bm{\Sigma}_{\widetilde{\bm{x}}^{(j)}}) = -\left(\nabla\nabla \log P(\bm{\widetilde{x}}^{(j)}|\ldots)\right |_{\bm{\widetilde{x}}^{(j)} = \bm{\mu}_{\widetilde{\bm{x}}^{(j)}}} )^{-1}$. The log full conditional distribution $h(\bm{\widetilde{x}}^{(j)}) = \log P(\bm{\widetilde{x}}^{(j)}|\ldots)$ is given by:
$$
h = \text{const} + \sum_{t=1}^{T} \bm{\widetilde{C}}^{'(j)}\left(\widetilde{\bm{y}}_t - \widetilde{\bm{\lambda}}_t\right) - \frac{1}{2}(\bm{\widetilde{x}}_1^{(j)}  - \bm{\widetilde{x}}_0^{(j)})'\bm{\widetilde{\bm{Q}}}^{(j)}_0(\bm{\widetilde{x}}_1^{(j)}  - \bm{\widetilde{x}}_0^{(j)}) - \sum_{t=2}^{T}\frac{1}{2}(\bm{\widetilde{x}}_t^{(j)}  - \widetilde{\bm{A}}^{(j)}\bm{\widetilde{x}}_{t-1}^{(j)})'\bm{\widetilde{\bm{Q}}}^{(j)}(\bm{\widetilde{x}}_t^{(j)}  - \widetilde{\bm{A}}^{(j)}\bm{\widetilde{x}}_{t-1}^{(j)})
$$

\textbf{It should be $\bm{b}_t^{(j)}$, but not $\bm{b}^{(j)}$} ! Fix that and then edit the writing!

\end{document}